---
title: "Final Project Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: bootstrap
    social: menu
    source_code: https://github.com/raleighgoodwin/EDLD-final-project
---

```{r setup, include=FALSE}
# Libraries

library(flexdashboard)
library(plotly)
library(shiny)
library(tidyverse)
library(here)
library(rio)
library(psych)
library(forcats)
library(scales)
library(lme4)
library(plyr)
library(dplyr)
library(colorspace)
library(lubridate)
library(ggtext)
library(gt)
# library(conflicted) # this worked at first and now R is mad at it and I don't know why
# 
# conflict_prefer("summarize", "dplyr")
# conflict_prefer("group_by", "dplyr")
```

```{r include=F}
# Data

dfc <- import(here::here("CCAMdata2008-2018.sav"))

dfc <- dfc %>% 
  mutate(yr = revalue(as.factor(year), c("1" = "2008",
                              "2" = "2010",
                              "3" = "2011",
                              "4" = "2012",
                              "5" = "2013",
                              "6" = "2014",
                              "7" = "2015",
                              "8" = "2016",
                              "9" = "2017",
                              "10" = "2018")),
         yr = as_date(yr))

```

```{r}
# Plot 1 work (setting up df)

dfc_voter <- dfc %>% 
  filter(registered_voter != 5 & party_w_leaners != -1 & wave > 6) %>% 
  mutate(registered_voter = factor(registered_voter,
                                   levels = c("1", "2", "3", "4")),
         party_w_leaners = as.factor(party_w_leaners)) %>% 
  na.omit()


```

```{r}
# Plot 2 work (setting up df)

dfc_harm <- dfc %>% 
  pivot_longer(
    cols = harm_personally:harm_plants_animals,
    names_to = "question",
    values_to = "harm"
  ) %>% 
  filter(harm != -1)

harm_plotL0 <- dfc_harm %>% 
  group_by(year, question) %>% 
  dplyr::summarise(tot = n()) %>% 
  slice(rep(1:n(), each = 5)) %>% 
  select(tot)

harm_plotR0 <- dfc_harm %>%
  group_by(year, question, harm) %>% 
  dplyr::summarise(n = n())

harm_plot0 <- cbind(harm_plotL0, harm_plotR0) %>% 
  select(-year...3) %>% 
  mutate(year = year...1,
         perc = (n / tot) * 100) %>% 
  select(year, question, harm, n, tot, perc)

dfc_harm <- dfc_harm %>% # making these values look nice for plot
  mutate(question2 = revalue(as.factor(question), 
                             c("harm_dev_countries" = "Developing Countries",
                               "harm_future_gen" = "Future Generations",
                               "harm_personally" = "You Personally",
                               "harm_plants_animals" = "Plants & Animals",
                               "harm_US" = "USA")),
         harm2 = revalue(as.factor(harm), 
                        c("0" = "Don't know",
                          "1" = "Not at all",
                          "2" = "Only a little",
                          "3" = "A moderate amount",
                          "4" = "A great deal")))

harm_plotL <- dfc_harm %>% 
  group_by(yr, question2) %>% 
  dplyr::summarise(tot = n()) %>% 
  slice(rep(1:n(), each = 5)) %>% 
  select(tot)

harm_plotR <- dfc_harm %>%
  group_by(yr, question2, harm2) %>% 
  dplyr::summarise(n = n())

harm_plot <- cbind(harm_plotL, harm_plotR) %>% 
  select(year = 1, everything(), -yr...3) %>% 
  mutate(perc = (n / tot) * 100,
         year = ymd(year, truncated = 2L)) %>% 
  select(year, question2, harm2, n, tot, perc) %>% 
  mutate(question2 = factor(question2, # making sure levels are in order I want for when I facet wrap
                            levels = c("You Personally", "Future Generations",
                                       "USA", "Developing Countries",
                                       "Plants & Animals")))
```

```{r}
# Plot 3 work (setting up df)

dfc_pol <- dfc %>% 
  pivot_longer(
    cols = reg_CO2_pollutant:reg_coal_emissions,
    names_to = "Policy",
    values_to = "Support"
  ) %>% 
  filter(Support != -1)

dfc_pol <- dfc_pol %>% # making these values look nice for plot
  mutate(Policy = revalue(as.factor(Policy), 
                             c("reg_CO2_pollutant" = "Regulate Carbon Dioxide Emissions",
                               "reg_utilities" = "Regulate Utility Energy Sources",
                               "fund_research" = "Fund Renewable Energy Research",
                               "reg_coal_emissions" = "Regulate Coal Power Plant Emissions")),
         Support = revalue(as.factor(Support), 
                        c("1" = "Strongly oppose",
                          "2" = "Somewhat oppose",
                          "3" = "Somewhat support",
                          "4" = "Strongly support")))

pol_plotL <- dfc_pol %>% 
  group_by(year, Policy) %>% 
  dplyr::summarise(totpol = n()) %>% 
  slice(rep(1:n(), each = 4)) %>% 
  select(totpol)

pol_plotR <- dfc_pol %>%
  group_by(year, Policy, Support) %>% 
  dplyr::summarise(npol = n())

pol_plot <- cbind(pol_plotL, pol_plotR) %>% 
  select(-year...3) %>% 
  mutate(year = year...1,
         percpol = (npol / totpol) * 100) %>% 
  select(year, Policy, Support, npol, totpol, percpol)

```

Data Source {.sidebar}
==========
**Data Source**

These figures utilize data from the [Climate Change in the American Mind (CCAM)](https://climatecommunication.yale.edu/visualizations-data/americans-climate-views/) project. Led by the [Yale Program on Climate Change Communication](https://climatecommunication.yale.edu/) and the [George Mason Center for Climate Change Communication](https://www.climatechangecommunication.org/), CCAM consists of national survey data on climate change beliefs collected between 2008 and 2018 (excluding 2009) and includes over 20,000 data points.

[Raleigh Goodwin](https://github.com/raleighgoodwin) created the plots on this page using the [ggplot](https://ggplot2.tidyverse.org/reference/ggplot.html) package.

**Citations:**

Yale Program on Climate Change Communication (YPCCC) & George Mason University Center for Climate Change Communication (Mason 4C). (2020). *Climate Change in the American Mind: National survey data on public opinion (2008-2018)* [Data file and codebook]. doi: [10.17605/OSF.IO/JW79P](https://osf.io/jw79p/)

Ballew, M. T., Leiserowitz, A., Roser-Renouf, C., Rosenthal, S. A., Kotcher, J. E., Marlon, J. R., Lyon, E., Goldberg, M. H., & Maibach, E. W. (2019). Climate Change in the American Mind: Data, tools, and trends. *Environment: Science and Policy for Sustainable Development, 61*(3), 4-18. doi: [10.1080/00139157.2019.1589300](https://www.tandfonline.com/doi/full/10.1080/00139157.2019.1589300)

Plot 1
==========

Column {data-width=700}
-----------------------------------------------------------------------

### **Final Version**

```{r}
dfc_voter %>% 
  dplyr::group_by(party_w_leaners, registered_voter) %>% 
  dplyr::summarize(mean_worry = mean(worry),
                   n = n(),
                   se = (sd(worry)/sqrt(n))) %>% 
  ggplot(aes(x = party_w_leaners, mean_worry, fill = as.factor(registered_voter))) +
  geom_col(position = "dodge",
           alpha = .85) +
  geom_errorbar(aes(ymin = mean_worry - se, ymax = mean_worry + se),
                width = .2,
                position = position_dodge(.9),
                color = "darkgray") +
  scale_fill_viridis_d(option = "plasma",
                       name = "Registered Voter",
                       labels = c("Yes", "No", "Don't know")) +
  scale_x_discrete(labels = c("Republican", "Democrat", "Independent/\nOther", "No party/\nApolitical")) +
  scale_y_continuous(labels = c("Not at all worried", "Not very worried", "Somewhat worried", "Very worried"),
                     breaks = 1:4, limits = c(0, 4)) +
  labs(y = "Average Worry", x = "Political Leanings",
       title = "Are Registered Voters More Concerned about Global Warming?",
       subtitle = 'Question: "How worried are you about global warming?"') +
  theme_light() +
  theme(plot.title.position = "plot",
        legend.position = "top",
        plot.subtitle = element_text(face = "bold"))

```

Column {.tabset data-width=350}
-----------------------------------------------------------------------

### About

In 2018, I ran a pilot study examining the effects of framing on pro-environmental behaviors, including donation to a pro-environmental organization. The results of this experience showed that only voter status predicted the amount participants were willing to donate. This figure shows a similar, yet more complex, relationship within the current data: The effect of voter registration on climate change concern is influenced by one's party affiliation. Simply put, among republicans, participants who did not know if they were registered to vote showed higher levels of concern regarding climate change than those who knew they were registered to vote. For democrats, independents, and those with no party affiliation, this relationship was opposite; those who did not know if they were registered to vote were less concerned. While democrats were most concerned about climate change, ov average, they only reported being "somewhat worried."

Error bars show +/- standard error. For information on the number of participants represented by each bar, see the table below.

```{r}
dfc_voter %>% 
  dplyr::group_by(party_w_leaners, registered_voter) %>% 
  dplyr::summarize(mean_worry = round(mean(worry), 2),
                   n = n()) %>% 
  mutate(registered_voter = revalue(as.factor(registered_voter), 
                                    c("1" = "Yes",
                                      "2" = "No",
                                      "4" = "Don't know")),
         party_w_leaners = revalue(as.factor(party_w_leaners), 
                                   c("1" = "Republican",
                                     "2" = "Democrat",
                                     "3" = "Independent/Other",
                                     "4" = "No party/Apolitical"))) %>% 
  select(party_w_leaners, registered_voter, n, mean_worry) %>% 
  gt() %>% 
  cols_label(registered_voter = "Registered Voter?",
               mean_worry = "Average Worry",
               n = "N") %>% 
  cols_align(align = "left",
             columns = vars(registered_voter)) %>% 
  tab_header(title = "Average Worry by Voter Registration and Political Affiliation")
  
```

### Version 1

This is the default plot generated by very basic ggplot code. For Version 2, I created meaningful labels and titles for the plot, changed the colorscheme to be colorblind friendly, and added error bars to facilitate comparisons of values. I also changed the theme for aesthetic purposes.

```{r}
dfc_voter %>% 
  dplyr::group_by(party_w_leaners, registered_voter) %>% 
  dplyr::summarize(mean_worry = mean(worry)) %>% 
  ggplot(aes(x = party_w_leaners, mean_worry, fill = as.factor(registered_voter))) +
  geom_col(position = "dodge")
```

### Version 2

For the final version, in order to allocate more space to the plot itself, I relocated the legend to the top of the plot. For aesthetics, I added transparency to the bars. For clarity, I created more meaningful labels for the y-axis and added a subtitle that displayed the survey question.

```{r}
dfc_voter %>% 
  dplyr::group_by(party_w_leaners, registered_voter) %>% 
  dplyr::summarize(mean_worry = mean(worry),
                   n = n(),
                   se = (sd(worry)/sqrt(n))) %>% 
  ggplot(aes(x = party_w_leaners, mean_worry, fill = as.factor(registered_voter))) +
  geom_col(position = "dodge") +
  geom_errorbar(aes(ymin = mean_worry - se, ymax = mean_worry + se),
                width = .2,
                position = position_dodge(.9)) +
  scale_fill_viridis_d(option = "plasma",
                       name = "Registered Voter",
                       labels = c("Yes", "No", "Don't know")) +
  scale_x_discrete(labels = c("Republican", "Democrat", "Independent/Other", "No party/Apolitical")) +
  labs(y = "Average Worry", x = "Political Leanings",
       title = '"How worried are you about global warming?"') +
  theme_light() +
  theme(plot.title.position = "plot")
```


Plot 2
==========

Column {data-width=700}
-----------------------------------------------------------------------

### **Final Version**

```{r}
harm_plot %>% 
  ggplot(aes(year, perc, color = as.factor(harm2))) +
  geom_line(size = .75) +
  facet_wrap(~question2,
             ncol = 2) +
  # scale_x_discrete(labels = xlab) +
  labs(x = "Year", y = "Percent of Responses",
       title = "Participants Largely Do Not Predict Personal Harm from Global Warming",
       subtitle = 'Question: "How much do you think global warming will harm..."') +
  scale_color_viridis_d(name = "Predicted Degree of Harm",
                        option = "plasma") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_light() +
  theme(plot.title.position = "plot",
        legend.position = c(.76, .08),
        plot.subtitle = element_text(face = "bold"))
```

Column {.tabset data-width=350}
-----------------------------------------------------------------------

### About

Participants may be reporting less concern because they erroneously do not percieve themselves to be at personal risk due to climate change. Over time, participants progressively predicted higher levels of risk for plants and animals, future generals, and developing countries. However, participants continuously underestimated the risk posed by climate change to themselves and their own country (USA). 

### Version 1

The first version of this plot is once again default ggplot output. For Version 2, I changed the theme and cleaned up labels for the axes, legend, and facets.

```{r}
harm_plot0 %>% 
  ggplot(aes(year, perc, color = as.factor(harm))) +
  geom_line() +
  facet_wrap(~question)
```

### Version 2

For the final version, I switched to a colorblind friendly palette, changed the theme again to match the other plots on my dashboard, and moved the legend into the empty space on the bottom right. I also added a title and subtitle to help with plot interpretation and percent signs to the y-axis labels.

```{r}
harm_plot %>% 
  ggplot(aes(year, perc, color = as.factor(harm2))) +
  geom_line() +
  facet_wrap(~question2) +
  labs(x = "Year",
       y = "Percent of Responses") +
  theme_minimal()
```


Plot 3
==========

Column {data-width=700}
-----------------------------------------------------------------------

### **Final Version**

```{r}

pol_plot %>%
  filter(year == 10) %>%
  ggplot(aes(Policy, percpol, fill = Support)) +
  geom_col(alpha = .85) +
  coord_flip() +
  scale_fill_viridis_d(option = "plasma",
                       guide = guide_legend(reverse = T)) +
  labs(y = "Percent of Responses",
       x = NULL,
       title = "Most Participants Supported Global Warming Mitigation Policies in 2018",
       subtitle = 'Question: "How much do you support or oppose..."\n\n') +
  scale_x_discrete(labels = c("Funding more research into\n renewable energy sources",
                              "Regulating CO[2] as a pollutant",
                              "Requiring electric utilities to\n produce at least 20% of electricity\n using renewable resources",
                              "Setting strict CO[2] emission\n limits on coal-fired power plants")) +
  theme_light() +
  theme(plot.title.position = "plot",
        legend.position = c(0,1),
        legend.justification = c(.36,0),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        axis.text.y = element_text(hjust = .5),
        plot.subtitle = element_text(face = "bold")) +
  scale_y_continuous(labels = scales::percent_format(scale = 1))

```

Column {.tabset data-width=350}
-----------------------------------------------------------------------

### About

Despite arguably underestimating the risk posed by climate change, participants in 2018 largely supported policies aimed at mitigating the effects of global warming. All four policies included in the survey recieved support from the majority of participants.

```{r}
pol_plot %>%
  filter(year == 10) %>% 
  select(Policy, Support, totpol, percpol) %>% 
  mutate(percpol = round(percpol, 0)/100,
         N = totpol) %>% 
  pivot_wider(names_from = Support, values_from = percpol) %>% 
  janitor::clean_names() %>% 
  select(policy, n, strongly_support, somewhat_support, somewhat_oppose, strongly_oppose) %>% 
  gt() %>% 
  fmt_percent(vars(strongly_support, somewhat_support, somewhat_oppose, strongly_oppose),
              decimals = 0) %>% 
  
  cols_label(policy = "Policy",
             n = "N",
             strongly_support = "Strongly Support",
             somewhat_support = "Somewhat Support",
             somewhat_oppose = "Somewhat Oppose",
             strongly_oppose = "Strongly Oppose") %>% 
  cols_align(align = "left",
             columns = vars(policy)) %>% 
  tab_header(title = "2018 Support for Four Climate Change Mitigation Policies")

```

### Version 1

This is once again the default ggplot output. For the second version, I switched to a colorblind friendly palette, fixed messy axis labels, and flipped the plot on its side to help with readability of the policy labels.

```{r}
pol_plot %>% 
  ggplot(aes(Policy, percpol, fill = Support)) +
  geom_col(position = "dodge")
```

### Version 2

For the third version, I fixed axes labels to incorporate percent signs and a more in-depth description of each policy. I also added a title to orient the audience on the survey question and changed the theme to match the other plots. Because the bars add to 100%, I stacked them to make comparisons between policies easier.

```{r}
pol_plot %>% 
  ggplot(aes(Policy, percpol, fill = Support)) +
  geom_col(position = "dodge") +
  coord_flip() +
  scale_fill_viridis_d(option = "plasma") +
  labs(y = "Percent of Responses")
```

### Version 3

For the final version of this plot, I added transparency to the bars, made the y-axis labels centered instead of right-aligned, and moved the legend to the top of the plot to give the plot itself more room. I also changed the title to aid in interpretation of the visual and added a subtitle that shows the survey question. 

```{r}
pol_plot %>%
  filter(year == 10) %>%
  ggplot(aes(Policy, percpol, fill = Support)) +
  geom_col() +
  coord_flip() +
  scale_fill_viridis_d(option = "plasma") +
  labs(y = "Percent of Responses",
       x = NULL,
       title = '"How much do you support or oppose..." (2018)') +
  scale_x_discrete(labels = c("Regulating CO^2 as a pollutant",
                              "Requiring electric utilities to\n produce at least 20% of electricity\n using renewable resources",
                              "Funding more research into\n renewable energy sources",
                              "Setting strict CO^2 emission\n limits on coal-fired power plants")) +
  theme_light() +
  scale_y_continuous(labels = scales::percent_format(scale = 1))
```
