---
title: "Final Project Dashboard"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: bootstrap
    social: menu
    source_code: https://raleighgoodwin.github.io/EDLD-final-project/
---

```{r setup, include=FALSE}
# Libraries

library(flexdashboard)
library(plotly)
library(shiny)
library(tidyverse)
library(here)
library(rio)
library(psych)
library(forcats)
library(scales)
library(lme4)
library(plyr)
library(dplyr)
library(colorspace)
library(lubridate)
```

```{r include=F}
# Data

dfc <- import(here::here("CCAMdata2008-2018.sav"))

dfc <- dfc %>% 
  mutate(yr = revalue(as.factor(year), c("1" = "2008",
                              "2" = "2010",
                              "3" = "2011",
                              "4" = "2012",
                              "5" = "2013",
                              "6" = "2014",
                              "7" = "2015",
                              "8" = "2016",
                              "9" = "2017",
                              "10" = "2018")),
         yr = as_date(yr))

```

```{r}
# Plot 1 work

dfc_harm <- dfc %>% 
  pivot_longer(
    cols = harm_personally:harm_plants_animals,
    names_to = "question",
    values_to = "harm"
  ) %>% 
  filter(harm != -1)

harm_plotL0 <- dfc_harm %>% 
  group_by(year, question) %>% 
  dplyr::summarise(tot = n()) %>% 
  slice(rep(1:n(), each = 5)) %>% 
  select(tot)

harm_plotR0 <- dfc_harm %>%
  group_by(year, question, harm) %>% 
  dplyr::summarise(n = n())

harm_plot0 <- cbind(harm_plotL0, harm_plotR0) %>% 
  select(-year...3) %>% 
  mutate(year = year...1,
         perc = (n / tot) * 100) %>% 
  select(year, question, harm, n, tot, perc)

dfc_harm <- dfc_harm %>% # making these values look nice for plot
  mutate(question2 = revalue(as.factor(question), 
                             c("harm_dev_countries" = "Developing Countries",
                               "harm_future_gen" = "Future Generations",
                               "harm_personally" = "You Personally",
                               "harm_plants_animals" = "Plants & Animals",
                               "harm_US" = "US")),
         harm2 = revalue(as.factor(harm), 
                        c("0" = "Don't know",
                          "1" = "Not at all",
                          "2" = "Only a little",
                          "3" = "A moderate amount",
                          "4" = "A great deal")))

harm_plotL <- dfc_harm %>% 
  group_by(year, question2) %>% 
  dplyr::summarise(tot = n()) %>% 
  slice(rep(1:n(), each = 5)) %>% 
  select(tot)

harm_plotR <- dfc_harm %>%
  group_by(year, question2, harm2) %>% 
  dplyr::summarise(n = n())

harm_plot <- cbind(harm_plotL, harm_plotR) %>% 
  select(-year...3) %>% 
  mutate(year = year...1,
         perc = (n / tot) * 100) %>% 
  select(year, question2, harm2, n, tot, perc) %>% 
  mutate(question2 = factor(question2, # making sure levels are in order I want for when I facet wrap
                            levels = c("You Personally", "Future Generations",
                                       "US", "Developing Countries",
                                       "Plants & Animals")))
```

```{r}
# Plot 2 work

dfc_pol <- dfc %>% 
  pivot_longer(
    cols = reg_CO2_pollutant:reg_coal_emissions,
    names_to = "Policy",
    values_to = "Support"
  ) %>% 
  filter(Support != -1)

dfc_pol <- dfc_pol %>% # making these values look nice for plot
  mutate(Policy = revalue(as.factor(Policy), 
                             c("reg_CO2_pollutant" = "Regulate Carbon Dioxide Emissions",
                               "reg_utilities" = "Regulate Utility Energy Sources",
                               "fund_research" = "Fund Renewable Energy Research",
                               "reg_coal_emissions" = "Regulate Coal Power Plant Emissions")),
         Support = revalue(as.factor(Support), 
                        c("1" = "Strongly oppose",
                          "2" = "Somewhat oppose",
                          "3" = "Somewhat support",
                          "4" = "Strongly support")))

pol_plotL <- dfc_pol %>% 
  group_by(year, Policy) %>% 
  dplyr::summarise(totpol = n()) %>% 
  slice(rep(1:n(), each = 4)) %>% 
  select(totpol)

pol_plotR <- dfc_pol %>%
  group_by(year, Policy, Support) %>% 
  dplyr::summarise(npol = n())

pol_plot <- cbind(pol_plotL, pol_plotR) %>% 
  select(-year...3) %>% 
  mutate(year = year...1,
         percpol = (npol / totpol) * 100) %>% 
  select(year, Policy, Support, npol, totpol, percpol)

```

```{r}
# Plot 3 work

dfc_voter <- dfc %>% 
  filter(registered_voter != 5 & party_w_leaners != -1 & wave > 6) %>% 
  mutate(registered_voter = factor(registered_voter,
                                   levels = c("1", "2", "3", "4")),
         party_w_leaners = as.factor(party_w_leaners)) %>% 
  na.omit()


```

Data Source {.sidebar}
==========
**Data Source**

These figures utilize data from the [Climate Change in the American Mind (CCAM)](https://climatecommunication.yale.edu/visualizations-data/americans-climate-views/) project. Led by the [Yale Program on Climate Change Communication](https://climatecommunication.yale.edu/) and the [George Mason Center for Climate Change Communication](https://www.climatechangecommunication.org/), CCAM consists of national survey data on climate change beliefs collected between 2008 and 2018 (excluding 2009) and includes over 20,000 data points.

[Raleigh Goodwin](https://github.com/raleighgoodwin) created the plots on this page using the [ggplot](https://ggplot2.tidyverse.org/reference/ggplot.html) package.

**Citations:**

Yale Program on Climate Change Communication (YPCCC) & George Mason University Center for Climate Change Communication (Mason 4C). (2020). *Climate Change in the American Mind: National survey data on public opinion (2008-2018)* [Data file and codebook]. doi: [10.17605/OSF.IO/JW79P](https://osf.io/jw79p/)

Ballew, M. T., Leiserowitz, A., Roser-Renouf, C., Rosenthal, S. A., Kotcher, J. E., Marlon, J. R., Lyon, E., Goldberg, M. H., & Maibach, E. W. (2019). Climate Change in the American Mind: Data, tools, and trends. *Environment: Science and Policy for Sustainable Development, 61*(3), 4-18. doi: [10.1080/00139157.2019.1589300](https://www.tandfonline.com/doi/full/10.1080/00139157.2019.1589300)

Background
==========

Plot 1
==========

Column {data-width=700}
-----------------------------------------------------------------------

### Current Version

```{r}
harm_plot %>% 
  ggplot(aes(year, perc, color = as.factor(harm2))) +
  geom_line(size = .75) +
  facet_wrap(~question2,
             ncol = 2) +
  # scale_x_discrete(labels = xlab) +
  labs(x = "Year", y = "Percent of Responses",
       title = '"How much do you think global warming will harm..."') +
  scale_color_viridis_d(name = "Predicted Degree of Harm",
                        option = "plasma") +
  theme_light() +
  theme(plot.title.position = "plot")
```

Column {.tabset data-width=350}
-----------------------------------------------------------------------

### Next Steps

Planning to fix the x axis text to accurately display the years. R has been giving me a lot of grief about this, and none of the fixes I've tried have worked.

### Version 1

```{r}
harm_plot0 %>% 
  ggplot(aes(year, perc, color = as.factor(harm))) +
  geom_line() +
  facet_wrap(~question)
```

### Version 2

```{r}
harm_plot %>% 
  ggplot(aes(year, perc, color = as.factor(harm2))) +
  geom_line() +
  facet_wrap(~question2) +
  theme_minimal()
```


Plot 2
==========

Column {data-width=700}
-----------------------------------------------------------------------

### Current Version

```{r}
pol_plot %>% 
  ggplot(aes(Policy, percpol, fill = Support)) +
  geom_col(position = "dodge") +
  coord_flip() +
  scale_fill_viridis_d(option = "plasma") +
  labs(y = "Percent of Responses",
       title = '"How much do you support or oppose..."') +
  theme_light() +
  theme(plot.title.position = "plot") +
  scale_x_discrete(labels = c("Regulating CO2 as a pollutant",
                              "Requiring electric utilities to produce\n  at least 20% of electricity using\n renewable resources",
                              "Funding more research into\n renewable energy sources",
                              "Setting strict CO2 emission\n limits on coal-fired power plants"))
```

Column {.tabset data-width=350}
-----------------------------------------------------------------------

### Next Steps

I'd like to model this plot similarly to the common Pew bar plot format, wherein the bars are stacked and all centered at the divide between support/oppose instead of being left aligned at zero. Plot also feels a bit busy/crowded right now.

### Version 1

```{r}
pol_plot %>% 
  ggplot(aes(Policy, percpol, fill = Support)) +
  geom_col(position = "dodge")
```

### Version 2

```{r}
pol_plot %>% 
  ggplot(aes(Policy, percpol, fill = Support)) +
  geom_col(position = "dodge") +
  coord_flip() +
  scale_fill_viridis_d(option = "plasma") +
  labs(y = "Percent of Responses")
```


Plot 3
==========

Column {data-width=700}
-----------------------------------------------------------------------

### Current Version

```{r}
dfc_voter %>% 
  dplyr::group_by(party_w_leaners, registered_voter) %>% 
  dplyr::summarize(mean_worry = mean(worry),
                   n = n(),
                   se = (sd(worry)/sqrt(n))) %>% 
  ggplot(aes(x = party_w_leaners, mean_worry, fill = as.factor(registered_voter))) +
  geom_col(position = "dodge") +
  geom_errorbar(aes(ymin = mean_worry - se, ymax = mean_worry + se),
                width = .2,
                position = position_dodge(.9)) +
  scale_fill_viridis_d(option = "plasma",
                       name = "Registered Voter",
                       labels = c("Yes", "No", "Don't know")) +
  scale_x_discrete(labels = c("Republican", "Democrat", "Independent/Other", "No party/Apolitical")) +
  scale_y_continuous(labels = c("Not at all worried", "Not very worried", "Somewhat worried", "Very worried"),
                     breaks = 1:4, limits = c(0, 4)) +
  labs(y = "Average Worry", x = "Political Leanings",
       title = '"How worried are you about global warming?"') +
  theme_light() +
  theme(plot.title.position = "plot")
```

Column {.tabset data-width=350}
-----------------------------------------------------------------------

### Next Steps

I'd like to put the legend at the bottom of the plot and fix the dimensions so that the x axis labels don't overlap. I'm also working on a related plot to this for an academic audience.

### Version 1

```{r}
dfc_voter %>% 
  dplyr::group_by(party_w_leaners, registered_voter) %>% 
  dplyr::summarize(mean_worry = mean(worry)) %>% 
  ggplot(aes(x = party_w_leaners, mean_worry, fill = as.factor(registered_voter))) +
  geom_col(position = "dodge")
```
